---
title: "Testing BCD Output"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---

Date: OCtober 1, 2019

Purpose: All the files have been written out for BCD.

## Summary

**The Good**:

1. The motifs being extracted are correct. For example the motifs that are scored above the threshold are all bcd motifs.  This is good!
2. The motif_extraction program is extracting all orthologous regions of each species, with the exception of when there are blanks in the align_positions of other species.

**The Bad**:

1. Whatever part of the map_extration program is choosing the what to extract is wrong.  I can not figure out how it is choosing these regions, but for sure it is not based on threshold. There are way too many. 
2. The marking of yes and no for motifs has nothing to do with being above the 7 threshold. 


## Part 1 - read in files and clean up a bit

Each region is a seperate file. 

```{r}
library(reshape2)
library(tidyverse)
library(ggplot2); theme_set(theme_bw())

f <- list.files("~/../../Volumes/ciera_2/2.re-running_whole_pipeline/output/bcd/")
length(f) 

```

Just a subset 
```{r, eval = FALSE}
## subset
file_list <- f[1:100]

## Get all files and put into a single list
## To do run with 500, it takes ~20 min 1:20


for (file in file_list){
       
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")){
    dataset <- read.csv(paste0("../../2.re-running_whole_pipeline/output/bcd/", file), header=TRUE)
  }
   
  # if the merged dataset does exist, append to it
  if (exists("dataset")){
    temp_dataset <-read.csv(paste0("../../2.re-running_whole_pipeline/output/bcd/", file), header=TRUE)
    dataset<-rbind(dataset, temp_dataset)
    rm(temp_dataset)
  }
 
}

dim(dataset)  

## Checkpoint
## write.csv(dataset, file = "../data/bcd_100_subset_20190903.csv")
```

Clean up the dataframe for better handling. Renaming ect.
```{r}
## dataset <- read.csv("../data/bcd_100_subset_20190903.csv")
dataset1 <- dataset

head(dataset1)
dim(dataset1)

## split ID to get region, enhancer_func, and species

dataset1 <- dataset1 %>%  
  separate(species_y, c("region", "enhancer_func", "species")) %>%
  select(-c("X"))

dataset1$raw_seq <- toupper(dataset1$raw_seq)
```

## Part 2: Motif Column

I will make a new motif column based on score and check.

```{r}

## Make a new column that labels "yes" to any score above 7
dataset1$new_motif <- ifelse(dataset1$score >= 7, "yes", "no")

## What it should look like.
dataset1 %>%
  ggplot(., aes(score, color = new_motif)) +
  geom_density()

## What it should look like for all the top motifs
dataset1 %>%
  filter(score >= 7.0) %>%
  group_by(raw_seq, new_motif) %>%
  tally() %>%
  ggplot(., aes(raw_seq,n, fill = new_motif)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Looking at one regions at a time

This is looking at just one region to see if I can figure out what is going on. 

````{r}
### Looking at the "yeses"
## Scores look correct when using the new_motif()
dataset1 %>% 
  filter(region == unique(dataset1$region)[6] & new_motif == "yes") %>%
  ggplot(., aes(align_position, species, color = score)) +
  geom_point() + scale_color_viridis_c() + facet_grid(.~strand)

dataset1 %>% 
  filter(region == unique(dataset1$region)[6]) %>%
  ggplot(., aes(align_position, species, color = score)) +
  geom_point() + scale_color_viridis_c() + facet_grid(.~strand)
```

When you look at everything that was extracted you see a few things.


```{r}
## This is what the map_motif extracted. 

dataset1 %>% 
  filter(region == unique(dataset1$region)[6]) %>%
  ggplot(., aes(align_position, species, color = score)) +
  geom_point() + scale_color_viridis_c() + facet_grid(.~strand)


## We should only have the
dataset1 %>% 
  filter(region == unique(dataset1$region)[10]) %>%
  ggplot(., aes(align_position, species, color = new_motif)) +
  geom_point() + facet_grid(.~strand)

```

## Part 3

Some work trying to quantify the motifs.

Start with asking - How many sites are shared by all species?

```{r}

dataset1 %>%
  group_by(align_position, region) %>%
  filter(new_motif  == "yes") %>%
  tally()
```

How many sites include all species?  Yes or no?

```{r}

## How can there be more than 24?
dataset1 %>%
  group_by(align_position, region, strand) %>%
  tally() %>%
  ggplot(., aes(n)) +
    geom_density()

## - [ ] Looks like it is all the same region, not sure what is going on here?
dataset1 %>%
  group_by(align_position, region, strand, raw_seq) %>%
  tally() %>%
  filter(n > 24)  %>%
  ggplot(., aes(n)) +
    geom_density()

dataset1 %>% 
  filter(region == "VT0849") %>%
  ggplot(., aes(align_position, species, color = new_motif)) +
  geom_point()  + facet_grid(.~strand)

```

I am just going to continue by subsetting above 24 till I figure out what is going on here.

While I am at it I am also going to remove all instances where there are only regions with less than two positions. 

```{r}
dataset1 %>%
  group_by(align_position, region, strand, new_motif, enhancer_func) %>%
  tally() %>%
  filter(n <= 24 & n > 2) %>%
  ggplot(., aes(n, color = new_motif)) +
    geom_density() +
    facet_grid(.~enhancer_func)

```

Is there are a difference for how many 
```{r}
dataset1 %>%
  group_by(enhancer_func,raw_seq) %>%
  tally() %>%
  arrange(-n) %>%
  top_n(40) %>%
  ggplot(., aes(raw_seq, n, fill = enhancer_func)) +
    geom_bar(stat = "identity", position = "dodge") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))


## How many motifs Avg in each region?
## There are just more TFBS per region
dataset1 %>%
  group_by(enhancer_func) %>%
  tally()

## Are the top motifs different between function and non-function?

dataset1 %>%
  group_by(enhancer_func,raw_seq) %>%
  filter(enhancer_func == 1) %>%
  tally() %>%
  top_n(40) %>%
  arrange(-n) %>%
  ggplot(., aes(reorder(raw_seq,-n), n)) +
    geom_bar(stat = "identity", position = "dodge") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Some questions I want to ask: 

1. Are these sites more conserved than surrounding regions?
  - Try just regions.
  - Get the simulated random sample control 
  - Get random region control
2. Look at positions and how they change


```{r}