---
output: html_document
editor_options: 
  chunk_output_type: console
---

What else needs to be one here?

We need to figure out why some regions have more than 24 species represented in each region.

https://www.biostars.org/p/76701/

- You need to combine 

```{r}

library(reshape2)
library(tidyverse)
library(ggplot2); theme_set(theme_bw())
library(reshape2)
library(googledrive)

## Read directly from Google drive
## file_list <- drive_ls(as_id("1fD_Ag5JZRtp1-4i_Qg9U76oPu5tmdti8"))
## drive_download(as_id(file_list$id[1]), overwrite = TRUE)
## read_csv("VT32774_final_raw.fa.csv")
  

```


## Part 0: Complile all data or just read in the compiled data

The below code reads in all the data and compiles into a single data file. This has been done previously, so all that is needed now is to read in the file. 

### Part 0 - Option 1: compile data
```{r}
## Uncomment to compile all data or just read in the data in the next cell
# directory <- "~/../../Volumes/ciera_2/2.re-running_whole_pipeline/data/output/bcd_2019_10_01/bcd/"
# 
# ## List files in directory for read in
# file_list <- list.files(directory)
# length(file_list)

# ## takes about 25 min
# for (file in file_list){
#        
#   # if the merged dataset doesn't exist, create it
#   if (!exists("dataset")){
#     dataset <- read.csv(paste0(directory, file), header=TRUE)
#     print("Does not exist")
#   }
#    
#   # if the merged dataset does exist, append to it
#   if (exists("dataset")){
#     temp_dataset <-read.csv(paste0(directory, file), header=TRUE)
#     dataset<-rbind(dataset, temp_dataset)
#     rm(temp_dataset)
#   }
#  
# }
# 
# dim(dataset)  
# head(dataset)
# write.csv(dataset, paste0(directory, "all_data_bcd_2019_10_01.csv"))

```

## Part 0 - Option 2: 

Read in the dataset from one datafile.
```{r}
dataset <- read.csv("~/../../Volumes/ciera_2/2.re-running_whole_pipeline/data/output/bcd_2019_10_01/bcd/all_data_bcd_2019_10_01.csv") 

## [ ] put into googledrive and download straight in.
```

```{r}
## Just so you can play with data inputed files without reading them in again
dataset1 <- dataset

## Ways to view and understand dataset
head(dataset1)
dim(dataset1)
str(dataset1)
colnames(dataset1)

head(dataset)
dataset1 %>% 
  separate(species_y, c("region", "enhancer_func", "species")) %>%
  select(-c("X")) 

dataset1 <- dataset1 %>% 
  separate(species_y, c("region", "enhancer_func", "species")) %>% 
 
  select(-c("X")) ##gets rid of column X

dataset1$TFBS_called <- ifelse(dataset1$score >= 7, "yes", "no")

head(dataset1)
```

## How many positions per region, per align_pos?

```{r}
## What the is going on here? Why are there more than 24?
dataset1$raw_seq <- toupper(dataset1$raw_seq)
dataset1 %>%
  group_by(region, align_position) %>%
  tally() %>%
  group_by(n) %>%
  tally() %>%
  ggplot(., aes(n, nn)) +
  geom_bar(stat = "identity") 
```

## This us cool, this means that there are about 15,000 regions that
## have all 24 species represented!

```{r}
dataset1 %>%
  group_by(region, align_position) %>%
  tally() %>%
  group_by(n) %>%
  tally() %>%  print(n=60)
  ggplot(., aes(n, nn)) +
  geom_bar(stat = "identity") + 
  xlim(0,26) +
  theme(text = element_text(size = 17),
        plot.title = element_text( size=14)) +
  labs(title="Represenative species at each alignment position \n", x = "number of species", y = "aligned motif regions") 
```

## How many times was the TFBS called

```{r}
## First we need to filter just the ones that have 24 represenative species
## all the regions that will be used to filter by 

dataset1 %>%
  group_by(region, align_position, strand) %>%
  tally() %>%
  filter(n == 24) 

### OMG I THINK THIS WORKED
### Reference: https://stackoverflow.com/questions/26573285/using-filter-with-count
only_24 <- dataset1 %>%
  group_by(region, align_position, strand) %>%
  filter(n() == 24)

## Test if it worked:
only_24 %>%
  group_by(region, align_position) %>%
  tally() %>%
  group_by(n) %>%
  tally() %>%  
  ggplot(., aes(n, nn)) +
  geom_bar(stat = "identity") + 
  xlim(0,26) +
  theme(text = element_text(size = 17),
        plot.title = element_text( size=14)) +
  labs(title="Represenative species at each alignment position \n", x = "number of species", y = "aligned motif regions") 

### Now everything will be a fraction of 24
only_24 %>%
  filter(TFBS_called == 'yes') %>%
  group_by(region, strand, align_position, TFBS_called) %>%
  tally() %>%
  group_by(n) %>%
  mutate(ratio = n/24) %>%
  group_by(ratio) %>%
  tally() %>%
  ggplot(., aes(ratio, n)) +
  geom_bar(stat = "identity") + 
  theme(text = element_text(size = 14),
        plot.title = element_text( size=14)) +
  labs(title="In each 24 species aligned motif set, \n what is the ratio bcd motifs? ", x = "ratio of bcd to non-bcd", y = "aligned motif regions") 
```

## What about overlap sights?

```{r}
rawseq <- dataset1 %>%
  group_by(raw_seq)
rawseq

#convert all factors to lowercase
rawseq$seq_lower = tolower(rawseq$raw_seq)
rawseq

#group by lowercase factors and sum over n
grouped_seq <- rawseq %>%
  group_by(seq_lower) %>% 
  mutate(n= n()) %>% 
     distinct(seq_lower, .keep_all=TRUE)

grouped_seq

#order by decreasing n
ordered_seq <- grouped_seq[with(grouped_seq, order(-n)), ]
ordered_seq

#create a column containing lists of bases
ordered_seq$split_seq = strsplit(ordered_seq$seq_lower, "")
ordered_seq

#assign a column to each base based on position
pos_seq <- ordered_seq %>% 
  separate(split_seq, c("pos0", "pos1", "pos2", "pos3", "pos4", "pos5", "pos6")) %>%
  select(-c("pos0"))
pos_seq

pos_seq_reshaped <- melt(pos_seq, measure.vars = c("pos1", "pos2", "pos3", "pos4", "pos5", "pos6"))

pos_seq_groups <- pos_seq_reshaped %>% group_by(enhancer_func, variable) %>%
             mutate(group_size = n()) %>%
             group_by(variable, value, enhancer_func) %>%
             summarise(perc = n()/max(group_size))

colnames(pos_seq_groups)[colnames(pos_seq_groups)=="variable"] <- "positions"
colnames(pos_seq_groups)[colnames(pos_seq_groups)=="value"] <- "nucleotides"
colnames(pos_seq_groups)[colnames(pos_seq_groups)=="perc"] <- "percentage"

pos_seq_groups

ggplot(data=pos_seq_groups, aes(fill=enhancer_func, y=percentage, x=nucleotides)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_y_continuous(labels = scales::percent) + 
  facet_grid(~positions)
```