---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


## Purpose

To map TFBS changes onto tree and visualize what is the prominant changes that are occuring. 

For those species that do, you can ask: How is this sequence different between them? Are these clade specific changes?  For those species that do not have the sequence, is this an instance of TFBS sequence decay?  How did this sequence decay?  It would be good to be able to identify, using the tree, what TFBS arose and which TFBS decayed. Still have to work out exactly how, but overall, we could use ancestral state reconstruction. 

We need the tree especially to create anscetral sequence reconstruction. Which I believe the best option is Phangorn as seen in this tutorial: [https://cran.r-project.org/web/packages/phangorn/vignettes/Ancestral.pdf](https://cran.r-project.org/web/packages/phangorn/vignettes/Ancestral.pdf).

We need to have the tree and redo the work that was done in [Moses et al., 2003](https://www.ncbi.nlm.nih.gov/pubmed/12946282).

```{r}

## Libraries
library(ggplot2)
library(reshape2)
library(stringr)
library(tidyverse)
library(ggtree)
library(skimr)
library(ape)
library(viridis)
library(viridisLite)
library(plyr)
#library(qdap)
library(stringr)
library(Biostrings)
#library(BSgenome)
#library(Rsamtools)
library(PWMEnrich)
library("colorspace")
```

```{r}
## Read in tree and clean
## Raw tree
tree <- read.tree("../data/tree/montium_phylogram_newick.tre")

## Before trimming
tree$tip.label

## Regions to remove
tips_remove <- c("DFengkainesis", "ananassae", "biarmipes","DTrapezifons","DParvula","DOrosa","DOgumai", "DOhnishii", "DLini","DMalagassya","DSeguyi", "DNikananu", "DBurlai", "elegans","biarmipes","simulans", "DGreeni", "DTruncata", "DTani", "DBicornuta", "DBocki", "triauraria_2_bronski", "DBarbarae", "DBocqueti", "DChauvacae", "DCurta_N", "DDiplacantha", "Drosophila_melanogaster")

## Remove regions
prunetree <- drop.tip(tree,tips_remove)

## Clean up names to match other data
prunetree$tip.label <- gsub("_bronski", "", prunetree$tip.label)
prunetree$tip.label <- paste0("D. ", prunetree$tip.label)

## Basic Bitch Tree
ggtree(prunetree) + 
  geom_text(aes(label = label), hjust=-.3) + 
  ggplot2::xlim(0, 0.07)

write.tree(prunetree, file = "../data/tree/cleanedTree.tre")

prunetree
```

## This is just an overview on how to visualize branches on a tree

```{r}
## Visualizing nodes
## Visualize

ggtree(prunetree) + 
  geom_text2(aes(#subset = isTip, # uncomment for just the tips
                 label = node), hjust = -.3) + 
  geom_tiplab(hjust = -.4) + 
  ggplot2::xlim(0, 0.07)

## This will be a good strategy.  We need to name by the nodes
cls <- list(c1 = c(22, 20, 33),
            c2 = 7)

tree <- groupOTU(prunetree, cls)

ggtree(tree, aes(color=group, linetype=group)) + geom_tiplab() +
     scale_color_manual(values=c("black", rainbow_hcl(4))) + theme(legend.position="right")

tree <- groupClade(prunetree, node=c(31,32, 35,36,37,42,43))
ggtree(tree, aes(color = group)) + 
  scale_color_manual(values = c("black", rep("pink",8)))
```


Below code is for matching to motif_e3xtract output
```{r, eval=FALSE}
## For switching out the species names in other files
species_key <- read.csv("../data/montium_species_laneID.csv")
species <- species_key$species
lane <- species_key$lane_ID

## pos
for(j in seq_along(lane)){
  pos$species <- gsub(lane[j], species[j], pos$species)
}

## occur
for(j in seq_along(lane)) {
  occur$species <- gsub(lane[j], species[j], occur$species)
}

#############################################
##  Making new column based on species number
############################################

occur$species_count  <- str_count(occur$species, 'D.')
```